On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-26 4_Result.tex
Updates from Overleaf

As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
code snippets, from which five were in their confusing versions, whilst the other five contained cleaned versions of the
code snippets (i.e., without the confusing constructs and idioms). 
The participants should provide the expected outcomes of the code snippets. As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). 
Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.

Considering Table~\ref{tab:difference-correctness}, the versions of six code snippets without atoms of confusion present at least a 15\% improvement in answer correctness. In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on
misunderstanding, and  
frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also introduce high degrees of confusion.
The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned version of the code snippets. Also, the sample of responses with no atoms had almost no dispersion, which supports the
argument that the non-confusing versions of the code snippets are
easier to evaluate correctly.

@@

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference
in the frequency of correct and incorrect answers---due to the versions
of the code snippets (confusing and clean code). The result indicates that the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$). We also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant
\emph{education} or participant \emph{experience} impact correctness. Although the results suggest that experience impacts correctness, the education level does not. As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.
Finally, we measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR)
method (see the Odds Ratio column of Table~\ref{tab:hypothesis-testing}). We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and
Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. For instance, the Comma Operator leads to an \emph{Odds Ratio} of
\num{19.02}, meaning that the the odds of a correct answer are \num{19.02} times
higher when interpreting the clean version of a code---in comparison with the corresponding, confusing version of the code snippet.

\begin{mh}
  The results of our exploratory data analysis and
  hypothesis testing suggest that the atom
  candidates Comma Operator, Post Increment,
  Omitted Curly Braces, Assignment as Value,
  Implicit Predicate, and Ternary Operator
  introduce some degree of misunderstanding
  in JavaScript code. 
\end{mh}

@@

 \centering
 \begin{tabular}{lcccc|cc}
   \toprule
Atom & ChiTest & Odds Ratio          & CI 2.5\% & CI 97.5\% & Wilcox Test & Cliff Delta \\ 
     &         & (misunderstanding)  &          &           & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & 6.60 & 68.22 & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & 9.21 & 356.62 & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & 1.73 & 15.74 & 0.12006 & 0.15 \\ 
Ommitted Curly Braces & 0.050962 & 2.35 & 1.00 & 5.77 & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & 1.81 & 79.12 & 0.82681 & 0.02 \\ 
Implict Predicate & 0.01123 & 6.95 & 1.46 & 66.56 & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & 0.73 & 3.28 & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & 0.74 & 12.57 & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & 0.55 & 2.85 & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & 0.22 & 3.12 & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
six atom candidates (Comma Operator, Automatic Semicolong Insertion,
Implicit Predicate, Logic
as Control Flow, Pre-Increment, and Arithmetic
as Logic). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%. 
For the first four, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting,
for the atom candidates Arithmetic as Logic and
Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method.
We found a large effect for the Comma Operator atom
candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

>>>>

As mentioned in Section~\ref{sec:meth:survey}, each participant in this study evaluated ten
code snippets, from which five were in their confusing versions, whilst the other five contained clean versions of the code snippets (i.e., without the confusing constructs and idioms). The participants should provide the expected outcomes of the code snippets. As discussed, we collected information about \emph{correctness} (whether the participant correctly predicted the program's output) and \emph{cognitive workload} (the time taken to answer the question). 
Table~\ref{tab:difference-correctness} and Figure~\ref{fig:boxplotcorrectness} summarize the results of the survey w.r.t. correctness.

Considering Table~\ref{tab:difference-correctness}, the clean versions of six code snippets present at least a 15\% improvement in answer correctness when compared with the confusing versions. In particular, the presence of the \emph{Comma Operator} atom presents the highest impact on misunderstanding, and frequently used constructs and idioms, such as \emph{Post Increment} and \emph{Omitted Curly Braces} (see Section~\ref{sec:msr-results}), also result in many mistakes.%introduce high degrees of confusion.
The boxplot of Figure~\ref{fig:boxplotcorrectness} shows a non-negligible decrease in the average number of incorrect answers when observing the cleaned versions of the code snippets. Also, the sample of responses with no atoms had almost no dispersion, which supports the
argument that the non-confusing versions of the code snippets are easier to evaluate correctly.

@@

We first use the \emph{Pearson's Chi-squared test}
to investigate if there is a statistically significant difference in the frequency of correct and incorrect answers---due to the versions of the code snippets (confusing and clean code). The p-values for these tests are reported in the ``Chi-square test'' column of Table~\ref{tab:hypothesis-testing}. The results indicate that for five atom candidates (Comma Operator, Automatic Semicolon Insertion, Post-Increment, Assignment as Value, Implicit Predicate) the confusing versions of the code snippets have a negative impact on code understanding (p-value $< 0.05$). This result holds even after applying the Benjamini-Hochberg correction with a false discovery rate of 5\%.

We measure the effect size of the clean version of the code snippets into the answers' correctness, using the \emph{Odds Ratio} (OR). The results are reported in the ``Odds Ratio'' column of Table~\ref{tab:hypothesis-testing}). In the table we also report the confidence interval (CI) for the OR in the ``CI'' column. Although many of the intervals are wide, for six of the atom candidates the lower bound of the confidence interval is greater than or equal to 1. This indicates that, at a 95\% confidence level, a developer is likely to commit less mistakes when using the clean versions.
%We found a negligible effect size for the atom candidates Arithmetic as Logic, Logic as Control Flow, and Pre-Increment. Nonetheless, for the remaining candidates, the effect size is significant. 
For instance, when comparing clean and confusing code snippets pertaining to the Comma Operator atom candidate, we observed an \emph{Odds Ratio} of \num{19.02}. This means that the odds of a correct answer are \num{19.02} times higher when interpreting the clean version---in comparison with the corresponding, confusing version of the code snippet. Furthermore, with a 95\% confidence level, the odds  ratio is between 6.6 and 68.22, i.e., although the error margin is wide, it is strongly in favor of the clean version. If there is a significant likelihood of a developer committing a mistake when analyzing a clean version, the lower bound of the CI should be a number between 0 and 1 (since the OR is a ratio). This is the case for the four atom candidates in the lower part of the table. In particular, in the last row (Arithmetic as Logic), the OR between 0 and 1 indicates that a study subject was actually more likely to misunderstand the clean version.

Finally, we also use the \emph{Binomial Generalized Logistic Regression} analysis to investigate if either participant \emph{education} or participant \emph{experience} impact correctness. Although the results suggest that experience impacts correctness, the education level does not. As such, we replicate the \emph{Pearson's Chi-squared test} for each experience group and found that the statistical difference is less significant for those developers with more than ten years of experience.

\begin{mh}
  The results of our exploratory data analysis and
  hypothesis testing suggest that the atom
  candidates Comma Operator, Automatic Semicolon Insertion, Post Increment, Omitted Curly Braces, 
  Assignment as Value, and
  Implicit Predicate %, and Ternary Operator
  introduce some degree of misunderstanding
  in JavaScript code. Based on the results of this study, they can be called atoms of confusion for JavaScript.
\end{mh}

@@

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom candidate & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Ommitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implict Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
six atom candidates (Comma Operator, Automatic Semicolong Insertion,
Implicit Predicate, Logic
as Control Flow, Pre-Increment, and Arithmetic
as Logic). For the first four, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

### 2021-04-26 5_Threats.tex
Updates from Overleaf

Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes. Threats to conclusion validity often come from inappropriate use of statistics. In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in empirical software engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient. For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these atoms are required.

@@

External validity is linked to whether it is possible to extrapolate the results of the study. Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are fairly frequent, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code.

>>>>

Conclusion validity is connected with how well it is possible to establish relationships between treatments and outcomes. Threats to conclusion validity often come from inappropriate use of statistics. In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient. For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these atoms are required.

@@

External validity is linked to whether it is possible to extrapolate the results of the study. Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
\castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

### 2021-04-26 main.tex
Updates from Overleaf

\newcommand{\na}{9\xspace}

@@

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) effort. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}6} out of \na atom candidates. In addition, our MSR effort reveals that atoms of confusion frequently arise in open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

>>>>

\newcommand{\na}{10\xspace}

@@

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}7} out of \na atom candidates. In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.
